\subsection{Shaltari Special Rules}

\subsubsection{Gateship-X}
Gates are special dropships used by Shaltari Fleets that can use teleportation networks. Gateships can deliver up to X gates per turn. Motherships may source gates by launching them, and gateships may source gates by redeploying them.
\\\\
When a gate is delivered to a gateship, it acts as a dropship that can target sites within 3" of the gateship. When a gate is sourced, it can be teleported through to any gateship to be delivered given that there is a chain of gateships, each within 18" of each other, connecting source and delivery ships.
\\\\
Gateships may redeploy up to X gates per turn, counting against the same limit as delivery. When a ground site is activated, before combat a player with gateships within dropship thrust range of the site can remove a friendly troop token to redeploy and deliver it elsewhere.\\\\
Gates can be launched to space stations or other orbital sites directly from a mothership if in range. Gateships are never considered to be launching, with only the Mothership gaining the associated spike.
\\\\
Gateships may only deliver or source gates when the Gateship has generated at least one power this turn.
\\\\
When an enemy troop is delivered to a surface site that is within 6" of a gateship in Atmosphere, a charged air roll may be made to destroy the troop before it lands. Roll a D6 for each troop, destorying the enemy if the TN 1 roll succeeds.

\subsubsection{Shaltari Shields}
When a Shaltari fleet Battlegroup activates, the controlling player may elect to raise or lower shields of each ship in the group. Ships with shields raised gain shield saves, but have an effective PD value of 0 and use their secondary signature value. Ships may not select the Silent Running special order while using shields.