\chapter{Core Concepts}

This chapter covers key concepts that are relevant to multiple parts of the game.

\section{Capital Ships}
Capital ships are any ships tonnage M, H, S or S2.

\section{Battlegroups and Squadrons}

Ships are added to fleets in \underline{Squadrons}, each of which occupies one fleet organization slot. Upon game start, all squadrons of the same ship class in a \underline{Battlegroup} fuse into a single squadron. Battlgroups are composed of one to three squadrons and each represented by a \underline{Strategy Card} which records the Strategy Rating and member squadrons of the Battlegroup.

\subsection{Strategy Rating}
The Strategy Rating on a Strategy card is set during fleet formation, and does not change as ships are destroyed during the game.

\subsection{Squadron Cohesion}
Squadrons must stay in cohesion in the game. A squadron is cohesive when at least one ship in the squadron is within coherency distance of all other ships in the squadron.
\endnote{This cohesion definition is somewhat easier to write, but it also means that squadrons can't spread out as much while still being slightly less cumbersome than 2.0 cohesion.}

\begin{itemize}
	\item L groups have a coherency distance of 3"
	\item Capital ships have a coherency distance of 6".
\end{itemize}

When moving their Squadrons, a player cannot voluntarily end a movement with ships out of cohesion.

\section{Heat and Signature}
The heat level of a ship affects how visible it is to enemies, and therefore how far away it can be targeted. A ships signature is the characteristic that forms the base of its \underline{Effective Signature}, which is a value based on the signature and heat level of the ship.

\begin{center}
	\begin{tabular}{|cc|}
		\hline
		\textbf{Heat} & \textbf{Effective Signature} \\
		\hline
		Silent Running &  0" \\
		\gray Normal & <Signature>"\\
		Minor & <Signature>" + 6" \\
		\gray Minor & <Signature>" + 12" \\
		\hline	
	\end{tabular} 
\end{center}
When a ship gains a spike, it moves down this table. When a ship removes a spike, it moves up this table, but cannot move to Silent Running except by use of the Silent Running special order. At all times the heat level of a ship must be tracked.

\section{Orbital Layers}
There are two orbital layers, orbit and atmosphere. The majority of gameplay will take place in orbit, while atmosphere hosts objectives and ground troops. At all times the orbital layer of a ship must be tracked.\\\\
IF a ship without the \underline{atmosphereic} rule would enter atmosphere, it is destroyed, and does not roll on the Catastrophic Damage table.

