\chapter{Troop Phase}

In the troop phase, weapons and other effects are activated and troops fight. In any order, activate each surface site and space station with troops on it. When activated, a surface site or space station performs the following steps:

\begin{enumerate}
	\item Feature Activation
	\item Troop Combat
	\item Resolution
\end{enumerate}

Troops are player aligned tokens that are placed on sites and stations. When all sites have activated, move to the launch phase.

\subsection{Feature Activation}
If a player has more troops on a site than all opponents, that player controls the site, and may use any weapons or other activated abilities of features on the site.

\subsection{Troop Combat}
In sites where opposing troops are present, each player rolls a D6 for each troop they control on the site. A players combat score is the highest die each player rolls. The opposing forces compare combat scores.
\begin{center}
	\resizebox{\textwidth}{!}{
 \begin{tabular}{c|c|c}
 	\hline
 	\textbf{Combat Scores} &  \textbf{Result Winner} & \textbf{Result Loser}  \\
 	\hline
 	Equal & \multicolumn{2}{c}{Players remove troops equally until only one player has troops left or all troops are removed.} \\
 	\hline
 	Not Equal & Remove all enemy troops & remove as many enemy troops as you have present, cannot remove last enemy troop. \\
 	\hline
 \end{tabular}}
\end{center}
 \endnote{This is 2.0 but with a bit of a twist that while still staying quick, will sometimes allow a player with fewer troops to win. As an example, in a 2v1 combat the outnumbered player has a 25\% chance of winning.}
 Apply troop losses to each side based on the  comparison of combat scores
 
 \subsection{Resolution}
 Any pending resolution effects are triggered.